### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9896714d2bd4a62ccff14475e979edd2152b1a35 🚀

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing SARS, MERS, and
COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.

>>>>

\vfill
Pathogen evolution is one of the least predictable components of disease
emergence, particularly in zoonotic reservoirs. Here, building on
principles established by the geographic mosaic theory of coevolution,
we develop a quantitative, spatially-explicit framework for mapping the
evolutionary risk of viral emergence. Driven by interest in diseases
like SARS, MERS, and COVID-19, we examine the global biogeography of
bat-origin betacoronaviruses, and apply our coevolutionary framework to
develop a map of emergence risk. We find that our framework helps
explain patterns like a unique pool of merbecoviruses in the Neotropics,
a recently-discovered lineage of divergent nobecoviruses in Madagascar,
and---most importantly---hotspots of diversification in southeast Asia,
sub-Saharan Africa, and the Middle East that correspond to the site of
previous zoonotic emergence events. Our framework may help identify
hotspots of future risk that have also been previously overlooked, like
west Africa and the Indian subcontinent, and may more broadly help
researchers understand how reservoir ecology shapes the evolution of
pandemic threats.

### 2022-06-28 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9896714d2bd4a62ccff14475e979edd2152b1a35 🚀

Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing SARS, MERS, and
COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.

>>>>

Pathogen evolution is one of the least predictable components of disease
emergence, particularly in zoonotic reservoirs. Here, building on
principles established by the geographic mosaic theory of coevolution,
we develop a quantitative, spatially-explicit framework for mapping the
evolutionary risk of viral emergence. Driven by interest in diseases
like SARS, MERS, and COVID-19, we examine the global biogeography of
bat-origin betacoronaviruses, and apply our coevolutionary framework to
develop a map of emergence risk. We find that our framework helps
explain patterns like a unique pool of merbecoviruses in the Neotropics,
a recently-discovered lineage of divergent nobecoviruses in Madagascar,
and---most importantly---hotspots of diversification in southeast Asia,
sub-Saharan Africa, and the Middle East that correspond to the site of
previous zoonotic emergence events. Our framework may help identify
hotspots of future risk that have also been previously overlooked, like
west Africa and the Indian subcontinent, and may more broadly help
researchers understand how reservoir ecology shapes the evolution of
pandemic threats.
